Trending assessment for nowcasts, measurements, and forecasts

### 2024-03-29 paper.tex
Some corrections; add short captions

\title{Trending assessment for nowcasts, measurements, and forecasts}
%\author{Oliver Grothe, Bolin Liu, Jonas Rieger}
\author[1]{Oliver Grothe}
\author[1]{Bolin Liu}
\author*[1]{Jonas Rieger}

@@

\maketitle
 \listoftodos

\section{Introduction}\label{sec:introduction}
The evaluation of measurement, prediction, and forecasting methods becomes increasingly important and sophisticated as technologies and data availability enable their application in more and more fields.
Conventionally, differences between predictions (or measurements) and target values, such as the \ac{rmse} or \ac{mae}, form the basis of method evaluation.
These measures do not consider trending, that is, whether the right direction of change is predicted or measured.
However, the information about an increase or decrease is crucial to detect changes in the underlying process early.
In the following, we provide a more specific overview of the characteristics and current evaluation schemes of the three application fields, forecasting, nowcasting, and measurement, and highlight why trending evaluation is highly relevant in the respective fields.

@@

\begin{table}
    \centering
    \scriptsize
    \begin{tabular}{l l}
        \toprule
        Application & Predicted change computation \\
        \midrule
        Measurement & $(x_{t} - x_{t-l})_{t=l}^T $\\
        Nowcasting & $
\diffxl =
\begin{cases}
(x_{t|t} - x_{t-l|t})^T_{t=l}, & \text{if } y_{t-l} \text{ is not known at time } t, \\
(x_{t|t} - y_{t-l})^T_{t=l}, & \text{otherwise}.
\end{cases}
$ \\
        Forecasting & $ \diffxl = (x_{t|t-l} - y_{t-l})^T_{t=l}$\\
        \bottomrule
    \end{tabular}
    \caption{Computation of the predicted change in the different applications. For nowcasting and forecasting, signals $x_{t | \tau}$ refer to values issued at $\tau$ with a target time $t$. For measurement, $x_t$ denotes the test device measurement at time $t$.}
    \label{tab:notation}
\end{table}

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{.24\textwidth}
\includegraphics{plots/illustrative_examples/4Q_sample_without_time}
\caption{Four-quadrant plot with simulated data.}\label{fig:trending_basic_4q_sample}
\end{subfigure}\hspace{0.01\textwidth}
\begin{subfigure}[t]{.24\textwidth}
\includegraphics{plots/illustrative_examples/4Q_sample_with_time}
\caption{The same data is colored according to the time index $t$, the greener, the later.}\label{fig:trending_basic_4q_sample_color}
\end{subfigure}\hspace{0.01\textwidth}
\begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/illustrative_examples/trending_ratio_time_series.pdf}
    \caption{Rolling estimate of the trending ratio over time with window length 91. }\label{fig:trending_ratio_time_series}
    \end{subfigure}%
    \caption{Visualizations for data with a time-varying trending ratio. We defer information on the data generation process to the appendix (see \ref{subsec:app-trending-data-generation}). The trending ratio for the entire data set is $\mu = 0.7577$. The strong seasonality of the trending ratio becomes visible in Figure~\ref{fig:trending_ratio_time_series}. The green curve $k_t$ shows the theoretical probability that $\diffxt$ has the same sign as $\diffyt$ for each time step. The rolling estimates are shifted to the right compared to $k_t$ as the windows look backward.}
\end{figure}

@@

\section{Application to measurement, nowcasting, and forecasting data} \label{sec:application}
In this section, we evaluate the trending in three different applications.
In the first application in Section~\ref{sec:application-covid}, we consider the nowcasting of the seven-day hospitalization rate for COVID-19 in Germany.
Section~\ref{sec:application-eda} applies the trending assessment to forecast methods for the number of arrivals in a large emergency department.
In the last application in Section~\ref{sec:application_measurement}, we consider the trending assessment of non-invasive blood pressure measurements compared to invasive blood pressure measurements.

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.48\textwidth}
    \includegraphics{plots/covid_nowcast/00_true_data.pdf}
    \caption{Realisations.}
    \label{fig:app-covid-true}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.48\textwidth}
    \includegraphics{plots/covid_nowcast/00_nowcast_data.pdf}
    \caption{Same-day nowcasts.}
    \label{fig:app-covid-nowcast}
        \end{subfigure}
    \caption{True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021, to April 29, 2022 \citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
    The outliers in the RKI model of values above $10^8$ are removed before the following analysis.}
    \label{fig:app-covid-true-nowcast}
\end{figure}

@@

\begin{table}[]
    \centering
    \input{plots/covid_nowcast/01_model_scores}
    \caption{Point evaluation measures for the issued mean of the different models. The evaluation period comprises 159 days. }
    \label{tab:app-covid-rmse}
\end{table}

@@

\begin{table}
    \centering
    \tiny
    \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
    \caption{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days. The exclusion areas are rectangles centered on the zero points with a width and height of twice the 10\%-quantile of the absolute values of nowcast and true values. The subscript $q_{0.1}$ denotes the measures with exclusion area. }
    \label{tab:app-covid-trending-ratios-lag-7}
\end{table}

@@

\begin{figure}
    \centering
%    \includegraphics{}
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_7}
    \caption{Conditional trending plot.}\label{fig:app-covid-cond-prob-7}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_7}
    \caption{Trending ratio over exclusion area size in $\diffx$.}\label{fig:app-covid-trending-ratio-7}
    \end{subfigure}
    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
    \label{fig:app-covid-cond-prob-trending-ratio-7}
\end{figure}

@@

\begin{table}
\centering
\input{plots/ed_arrival/00_point_evaluation_measures}
\caption{Point evaluation measures for the models. The smaller count for some models stems from missing forecasts scattered throughout the evaluation period.}\label{tab:app-eda-point-evaluation}
\end{table}

@@

\begin{table}
    \centering
    \tiny
    \input{plots/ed_arrival/10_marginal_analysis}
    \caption{Marginal analysis of the forecast and true differences. The column (1) shows the fraction of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values. Column (2) shows the fraction of values not in the exclusion area of size one.}
    \label{tab:app-eda-marginals}
\end{table}

@@

\begin{table}
    \centering
    \tiny
    \input{plots/ed_arrival/50_trending_ratio}
    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with the exclusion of zero-containing points for the horizons 72 hours and seven days.}
    \label{tab:app-eda-trending-ratios}
\end{table}

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.48\textwidth}
    \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_3}
    \caption{Horizon three days}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.48\textwidth}
    \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_7}
    \caption{Horizon seven days}
    \end{subfigure}
    \caption{Conditional trending plots for the horizons three and seven days and the models with the best positive or negative trending ability. The plots of NBI-2 and Poisson-2 are indistinguishable.}
    \label{fig:app-eda-cond-prob}
\end{figure}

\begin{figure}
    \begin{subfigure}{0.32\textwidth}
    \tiny
    \input{plots/ed_arrival/70_brier_scores}
    \caption{BS for the different models and horizons.}\label{fig:app-eda-prob-brier}
    \end{subfigure}\hspace{0.01\textwidth}%
    \begin{subfigure}[t]{0.32\textwidth}
    \includegraphics{plots/ed_arrival/60_reliability_diagram_lag_3}
    \caption{Reliability diagram for horizon three days.}\label{fig:app-eda-prob-rel-3}
    \end{subfigure}\hspace{0.01\textwidth}%
    \begin{subfigure}[t]{0.32\textwidth}
    \includegraphics{plots/ed_arrival/60_reliability_diagram_lag_7}
    \caption{Reliability diagram for horizon seven days.}\label{fig:app-eda-prob-rel-7}
    \end{subfigure}
    \caption{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days. The \acl{bs} in Figure~\ref{fig:app-eda-prob-brier} evaluates the calibration and sharpness of the probabilistic trending simultaneously, while the two plots on the right assess solely the calibration, that is, whether the predicted probability of increase occurs empirically. Probabilistic trending by the \ac{bs} is best for the models NBI-2 and Poisson-2 for both lags.}
    \label{fig:app-eda-prob}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics{plots/mimic/plot_4q}
    \caption{Four-quadrant plots for the different horizons and the systolic and mean blood pressure measurements. The upper row contains systolic measurements, and the lower row contains mean measurements. The columns contain the horizons one, five, and 15 minutes.}
    \label{fig:app-mimic-4q}
\end{figure}

@@

\section{Discussion and Conclusion}\label{sec:conclusion}

@@

\begin{figure}
    \centering
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/illustrative_examples/appendix_4q_dgp1}
        \caption{Constant trending ratio.}
    \end{subfigure}\hspace{0.01\textwidth}
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/illustrative_examples/appendix_4q_dgp1_time}
        \caption{Time-varying trending ratio}
    \end{subfigure}\hspace{0.01\textwidth}
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/illustrative_examples/appendix_4q_dgp1_asym}
        \caption{Asymmetric trending ratio}
    \end{subfigure}\hspace{0.01\textwidth}
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/illustrative_examples/appendix_4q_dgp2}
        \caption{Second approach}
    \end{subfigure}
    \caption{Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{subsec:app-trending-data-generation}. Although the first and second plots differ over time, their difference is not discernible in the plots. The third data set's asymmetry is visible in the plot, but the decrease in the trending ability near 0 is not visible. }
    \label{fig:appendix_dgps}
\end{figure}

@@

\begin{table}
    \centering
    \tiny
    \begin{subtable}{.48\textwidth}
        \input{plots/illustrative_examples/bootstrap_butterfly_cis}
        \caption{First dataset}
    \end{subtable}\hspace{0.02\textwidth}
    \begin{subtable}{.48\textwidth}
        \input{plots/illustrative_examples/bootstrap_normal_cis}
        \caption{Second dataset}
    \end{subtable}
    \caption{Proportion of bootstrapping confidence intervals covering the true value of trending ratio per method and sample size $T$. The average width of the confidence interval is listed in brackets.}
    \label{tab:trending_bootstrap}
\end{table}

\begin{figure}
    \centering
    \begin{subfigure}{0.48\textwidth}
\includegraphics{plots/illustrative_examples/boxplot_comp_time_butterfly}
        \caption{First dataset}
    \end{subfigure}
    \begin{subfigure}{0.48\textwidth}
    \includegraphics{plots/illustrative_examples/boxplot_comp_time_normal}
        \caption{Second dataset}
    \end{subfigure}
    \caption{Boxplot of the computation time of the different bootstrapping methods and data set sizes $T$. The computation time refers to bootstrapping one confidence interval based upon $10,000$ values. Each boxplot reflects $10,000$ samples. The \ac{bca} method takes slightly longer than the other two, but the difference is negligible.}
    \label{fig:trending_bootstrap_time}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}{.48\textwidth}
        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_butterfly}
        \caption{First dataset with asymmetric dependence.}
    \end{subfigure}
    \begin{subfigure}{.48\textwidth}
        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_normal}
        \caption{Second dataset. }
    \end{subfigure}
    \caption{Conditional trending plot for different bandwidth selection processes. Cross-validation least squares takes a considerably larger computation time. It does not converge for the first data set and yields a bandwidth too small for the second data set. The rule of thumb is the fastest method but tends to oversmooth. The cross-validation maximum likelihood method yields a more reasonable bandwidth with moderate computation time. $\varepsilon$ specifies an exclusion area $E = \{(x, y) \in \R^2: (-\varepsilon \leq x \leq \varepsilon)\}$ in $\diffx$-direction.}\label{fig:trending-cond-prob-bw}
\end{figure}

@@

\begin{table}
    \centering
    \begin{tabular}{l l}
        \toprule
        Abbreviation & Nowcasting hub key \\
        \midrule
        EPI & Epiforecasts-independent \\
        ILM & ILM-prop \\
        KIT & KIT-simple\_nowcast \\
        LMU & LMU\_StaBLab-GAM\_nowcast \\
        RIVM & RIVM-KEW \\
        RKI & RKI-weekly\_report \\
        SU & SU-hier\_bayes \\
        SZ & SZ-hosp\_nowcast\\
        ENS-MEAN & NowcastHub-MeanEnsemble\\
        ENS-MED & NowcastHub-MedianEnsemble\\
        \bottomrule
    \end{tabular}
    \caption{Matching the abbreviation to the key in the nowcasting hub.
    Information on the models and references is listed in \citet[][Table 1]{Wolffram2023}.}
    \label{tab:app-covid-models}
\end{table}

\begin{table}
    \centering
    \tiny
    \input{plots/covid_nowcast/10_marginal_analysis}
    \caption{Analysis of the nowcast and true differences for the horizons 1, 7, and 14 days.
    The column (1), $l=l$ shows the number of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
    \label{tab:app-covid-marginals}
\end{table}

\begin{table}
    \centering
    \begin{subtable}[t]{\textwidth}
        \input{plots/covid_nowcast/30_trending_ratios_lag_1}
    \caption{1 day.}
    \end{subtable}
    \begin{subtable}[t]{\textwidth}
        \input{plots/covid_nowcast/30_trending_ratios_lag_14}
        \caption{14 days.}
    \end{subtable}
    \caption{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the horizon 1 and 14 days. The exclusion areas are rectangles centered on the zero points with a width and height of 10\% of the quantile of the absolute values of nowcast and true values. }
    \label{tab:app-covid-trending-ratios-lag-1-14}
\end{table}

\begin{figure}
\centering
\includegraphics{plots/covid_nowcast/30_4q_plots}
\caption{Four-quadrant plots for the models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days. The spread in both directions increases with the horizon. }
\label{fig:app-covid-4q}
\end{figure}

\begin{figure}
    \centering
%    \includegraphics{}
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_1}
    \caption{Conditional trending plot for horzion 1.}\label{fig:app-covid-cond-prob-1}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_14}
    \caption{Conditional trending plot for horizon 1.}\label{fig:app-covid-cond-prob-14}
    \end{subfigure}
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_1}
    \caption{Trending ratio over exclusion area size in $\diffx$ for horizon 1.}\label{fig:app-covid-trending-ratio-1}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_14}
    \caption{Trending ratio over exclusion area size in $\diffx$ for horizon 1.}\label{fig:app-covid-trending-ratio-14}
    \end{subfigure}
    \caption{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the lag seven days.}
    \label{fig:app-covid-cond-prob-trending-ratio-1-14}
\end{figure}

>>>>

\title{Trending assessment for nowcasts, measurements, and forecasts}
%\author{Oliver Grothe, Bolin Liu, Jonas Rieger}
\author*[1]{Jonas Rieger}
\author[1]{Bolin Liu}
\author[1]{Oliver Grothe}

@@

\maketitle
% \listoftodos

\section{Introduction}\label{sec:trending-introduction}

The evaluation of measurement, prediction, and forecasting methods becomes increasingly important and sophisticated as technologies and data availability enable their application in more and more fields.
Conventionally, differences between predictions (or measurements) and target values, such as the \ac{rmse} or \ac{mae}, form the basis of method evaluation.
These measures do not consider trending, that is, whether the right direction of change is predicted or measured.
However, the information about an increase or decrease is crucial to detect changes in the underlying process early.
In the following, we provide a more specific overview of the characteristics and current evaluation schemes of the three application fields, forecasting, nowcasting, and measurement, and highlight why trending evaluation is highly relevant in the respective fields.

@@

\begin{table}
    \centering
    \scriptsize
    \begin{tabular}{l l}
        \toprule
        Application & Predicted change computation \\
        \midrule
        Measurement & $(x_{t} - x_{t-l})_{t=l}^T $\\
        Nowcasting & $
\diffxl =
\begin{cases}
(x_{t|t} - x_{t-l|t})^T_{t=l}, & \text{if } y_{t-l} \text{ is not known at time } t, \\
(x_{t|t} - y_{t-l})^T_{t=l}, & \text{otherwise}.
\end{cases}
$ \\
        Forecasting & $ \diffxl = (x_{t|t-l} - y_{t-l})^T_{t=l}$\\
        \bottomrule
    \end{tabular}
    \caption[Computation of the predicted change in the different applications.]{Computation of the predicted change in the different applications. For nowcasting and forecasting, signals $x_{t | \tau}$ refer to values issued at $\tau$ with a target time $t$. For measurement, $x_t$ denotes the test device measurement at time $t$.}
    \label{tab:notation}
\end{table}

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{.24\textwidth}
\includegraphics{plots/illustrative_examples/4Q_sample_without_time}
\caption{Four-quadrant plot with simulated data.}\label{fig:trending_basic_4q_sample}
\end{subfigure}\hspace{0.01\textwidth}
\begin{subfigure}[t]{.24\textwidth}
\includegraphics{plots/illustrative_examples/4Q_sample_with_time}
\caption{The same data is colored according to the time index $t$, the greener, the later.}\label{fig:trending_basic_4q_sample_color}
\end{subfigure}\hspace{0.01\textwidth}
\begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/illustrative_examples/trending_ratio_time_series.pdf}
    \caption{Rolling estimate of the trending ratio over time with window length 91. }\label{fig:trending_ratio_time_series}
    \end{subfigure}%
    \caption[Visualizations of the four-quadrant plot for data with a time-varying trending ratio.]{Visualizations for data with a time-varying trending ratio. We defer information on the data generation process to the appendix (see \ref{subsec:app-trending-data-generation}). The trending ratio for the entire data set is $\mu = 0.7577$. The strong seasonality of the trending ratio becomes visible in Figure~\ref{fig:trending_ratio_time_series}. The green curve $k_t$ shows the theoretical probability that $\diffxt$ has the same sign as $\diffyt$ for each time step. The rolling estimates are shifted to the right compared to $k_t$ as the windows look backward.}
\end{figure}

@@

\section{Application to measurement, nowcasting, and forecasting data} \label{sec:application}

In this section, we evaluate the trending in three different applications.
In the first application in Section~\ref{sec:application-covid}, we consider the nowcasting of the seven-day hospitalization rate for COVID-19 in Germany.
Section~\ref{sec:application-eda} applies the trending assessment to forecast methods for the number of arrivals in a large emergency department.
In the last application in Section~\ref{sec:application_measurement}, we consider the trending assessment of non-invasive blood pressure measurements compared to invasive blood pressure measurements.

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.48\textwidth}
    \includegraphics{plots/covid_nowcast/00_true_data.pdf}
    \caption{Realisations.}
    \label{fig:app-covid-true}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.48\textwidth}
    \includegraphics{plots/covid_nowcast/00_nowcast_data.pdf}
    \caption{Same-day nowcasts.}
    \label{fig:app-covid-nowcast}
        \end{subfigure}
    \caption[True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021, to April 29, 2022.]{True and nowcast data of the seven-day-hospitalization in Germany from November 22, 2021, to April 29, 2022 \citep{ChairOfEconometricsAndStatisticsAtKarlsruheInstituteOfTechnology2024}.
    The outliers in the RKI model of values above $10^8$ are removed before the following analysis.}
    \label{fig:app-covid-true-nowcast}
\end{figure}

@@

\begin{table}[]
    \centering
    \input{plots/covid_nowcast/01_model_scores}
    \caption[The trending ratios for Covid nowcasting.]{Point evaluation measures for the issued mean of the different models. The evaluation period comprises 159 days. }
    \label{tab:app-covid-rmse}
\end{table}

@@

\begin{table}
    \centering
    \tiny
    \input{plots/covid_nowcast/30_trending_ratios_lag_7.tex}
    \caption[The trending ratios for Covid nowcasting.]{The trending ratio $\accl[7]$, positive trending ratio $\accpl[7]$, and negative trending ratio $\accml[7]$ for the models without and with exclusion areas for the horizon seven days. The exclusion areas are rectangles centered on the zero points with a width and height of twice the 10\%-quantile of the absolute values of nowcast and true values. The subscript $q_{0.1}$ denotes the measures with exclusion area. }
    \label{tab:app-covid-trending-ratios-lag-7}
\end{table}

@@

\begin{figure}
    \centering
%    \includegraphics{}
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_cond_prob_lag_7}
    \caption{Conditional trending plot.}\label{fig:app-covid-cond-prob-7}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{.48\textwidth}
    \includegraphics{plots/covid_nowcast/40_acc_eps_lag_7}
    \caption{Trending ratio over exclusion area size in $\diffx$.}\label{fig:app-covid-trending-ratio-7}
    \end{subfigure}
    \caption[Conditional trending plot and trending ratio over exclusion area for the Covid nowcasts.]{Conditional trending plot and trending ratio over exclusion area for the nowcasts of the seven-day hospitalization rate ILM, RKI, RIVM, and ENS-MED for the horizon seven days.}
    \label{fig:app-covid-cond-prob-trending-ratio-7}
\end{figure}

@@

\begin{table}
\centering
\input{plots/ed_arrival/00_point_evaluation_measures}
\caption[Point evaluation measures for the emergency department arrival forecasting models.]{Point evaluation measures for the models. The smaller count for some models stems from missing forecasts scattered throughout the evaluation period.}\label{tab:app-eda-point-evaluation}
\end{table}

@@

\begin{table}
    \centering
    \tiny
    \input{plots/ed_arrival/10_marginal_analysis}
    \caption[Marginal analysis of the forecast and true differences in ermergency department arrival forecasting.]{Marginal analysis of the forecast and true differences. The column (1) shows the fraction of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values. Column (2) shows the fraction of values not in the exclusion area of size one.}
    \label{tab:app-eda-marginals}
\end{table}

@@

\begin{table}
    \centering
    \tiny
    \input{plots/ed_arrival/50_trending_ratio}
    \caption[Trending ratios for the emergency deparment arrival forecasts.]{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with the exclusion of zero-containing points for the horizons 72 hours and seven days.}
    \label{tab:app-eda-trending-ratios}
\end{table}

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.48\textwidth}
    \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_3}
    \caption{Horizon three days}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.48\textwidth}
    \includegraphics{plots/ed_arrival/50_Cond_Prob_lag_7}
    \caption{Horizon seven days}
    \end{subfigure}
    \caption[Conditional trending plots for the horizons three and seven days and the models with the best positive or negative trending ability in emergency department arrival forecasting.]{Conditional trending plots for the horizons three and seven days and the models with the best positive or negative trending ability. The plots of NBI-2 and Poisson-2 are indistinguishable.}
    \label{fig:app-eda-cond-prob}
\end{figure}

\begin{figure}
    \begin{subfigure}{0.32\textwidth}
    \tiny
    \input{plots/ed_arrival/70_brier_scores}
    \caption{BS for the different models and horizons.}\label{fig:app-eda-prob-brier}
    \end{subfigure}\hspace{0.01\textwidth}%
    \begin{subfigure}[t]{0.32\textwidth}
    \includegraphics{plots/ed_arrival/60_reliability_diagram_lag_3}
    \caption{Reliability diagram for horizon three days.}\label{fig:app-eda-prob-rel-3}
    \end{subfigure}\hspace{0.01\textwidth}%
    \begin{subfigure}[t]{0.32\textwidth}
    \includegraphics{plots/ed_arrival/60_reliability_diagram_lag_7}
    \caption{Reliability diagram for horizon seven days.}\label{fig:app-eda-prob-rel-7}
    \end{subfigure}
    \caption[Probabilistic trending evaluation in emergency department arrival forecasting.]{Probabilistic trending evaluation for the models Benchmark-1, GBM-2, NBI-2, Poisson-2, and qreg-1 for the horizons three and seven days. The \acl{bs} in Figure~\ref{fig:app-eda-prob-brier} evaluates the calibration and sharpness of the probabilistic trending simultaneously, while the two plots on the right assess solely the calibration, that is, whether the predicted probability of increase occurs empirically. Probabilistic trending by the \ac{bs} is best for the models NBI-2 and Poisson-2 for both lags.}
    \label{fig:app-eda-prob}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics{plots/mimic/plot_4q}
    \caption[Four-quadrant plots for the different horizons and the systolic and mean blood pressure measurements. ]{Four-quadrant plots for the different horizons and the systolic and mean blood pressure measurements. The upper row contains systolic measurements, and the lower row contains mean measurements. The columns contain the horizons one, five, and 15 minutes.}
    \label{fig:app-mimic-4q}
\end{figure}

@@

\section{Discussion and Conclusion}\label{sec:trending-conclusion}

@@

\begin{figure}
    \centering
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/illustrative_examples/appendix_4q_dgp1}
        \caption{Constant trending ratio.}
    \end{subfigure}\hspace{0.01\textwidth}
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/illustrative_examples/appendix_4q_dgp1_time}
        \caption{Time-varying trending ratio}
    \end{subfigure}\hspace{0.01\textwidth}
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/illustrative_examples/appendix_4q_dgp1_asym}
        \caption{Asymmetric trending ratio}
    \end{subfigure}\hspace{0.01\textwidth}
    \begin{subfigure}{0.24\textwidth}
        \includegraphics{plots/illustrative_examples/appendix_4q_dgp2}
        \caption{Second approach}
    \end{subfigure}
    \caption[Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{subsec:app-trending-data-generation}.]{Four-quadrant plots for sample realizations of the data generation schemes of Section~\ref{subsec:app-trending-data-generation}. Although the first and second plots differ over time, their difference is not discernible in the plots. The third data set's asymmetry is visible in the plot, but the decrease in the trending ability near 0 is not visible. }
    \label{fig:appendix_dgps}
\end{figure}

@@

\begin{table}
    \centering
    \tiny
    \begin{subtable}{.48\textwidth}
        \input{plots/illustrative_examples/bootstrap_butterfly_cis}
        \caption{First dataset}
    \end{subtable}\hspace{0.02\textwidth}
    \begin{subtable}{.48\textwidth}
        \input{plots/illustrative_examples/bootstrap_normal_cis}
        \caption{Second dataset}
    \end{subtable}
    \caption[Proportion of bootstrapping confidence intervals covering the true value of trending ratio per method and sample size $T$ in the bootstrap simulation study.]{Proportion of bootstrapping confidence intervals covering the true value of trending ratio per method and sample size $T$. The average width of the confidence interval is listed in brackets.}
    \label{tab:trending_bootstrap}
\end{table}

\begin{figure}
    \centering
    \begin{subfigure}{0.48\textwidth}
\includegraphics{plots/illustrative_examples/boxplot_comp_time_butterfly}
        \caption{First dataset}
    \end{subfigure}
    \begin{subfigure}{0.48\textwidth}
    \includegraphics{plots/illustrative_examples/boxplot_comp_time_normal}
        \caption{Second dataset}
    \end{subfigure}
    \caption[Boxplot of the computation time of the different bootstrapping methods and data set sizes $T$.]{Boxplot of the computation time of the different bootstrapping methods and data set sizes $T$. The computation time refers to bootstrapping one confidence interval based upon $10,000$ values. Each boxplot reflects $10,000$ samples. The \ac{bca} method takes slightly longer than the other two, but the difference is negligible.}
    \label{fig:trending_bootstrap_time}
\end{figure}

@@

\begin{figure}
    \centering
    \begin{subfigure}{.48\textwidth}
        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_butterfly}
        \caption{First dataset with asymmetric dependence.}
    \end{subfigure}
    \begin{subfigure}{.48\textwidth}
        \includegraphics{plots/illustrative_examples/cond_prob_plot_bw_normal}
        \caption{Second dataset. }
    \end{subfigure}
    \caption[Conditional trending plots for different bandwidth selection processes.]{Conditional trending plot for different bandwidth selection processes. Cross-validation least squares takes a considerably larger computation time. It does not converge for the first data set and yields a bandwidth too small for the second data set. The rule of thumb is the fastest method but tends to oversmooth. The cross-validation maximum likelihood method yields a more reasonable bandwidth with moderate computation time. $\varepsilon$ specifies an exclusion area $E = \{(x, y) \in \R^2: (-\varepsilon \leq x \leq \varepsilon)\}$ in $\diffx$-direction.}\label{fig:trending-cond-prob-bw}
\end{figure}

@@

\begin{table}
    \centering
    \begin{tabular}{l l}
        \toprule
        Abbreviation & Nowcasting hub key \\
        \midrule
        EPI & Epiforecasts-independent \\
        ILM & ILM-prop \\
        KIT & KIT-simple\_nowcast \\
        LMU & LMU\_StaBLab-GAM\_nowcast \\
        RIVM & RIVM-KEW \\
        RKI & RKI-weekly\_report \\
        SU & SU-hier\_bayes \\
        SZ & SZ-hosp\_nowcast\\
        ENS-MEAN & NowcastHub-MeanEnsemble\\
        ENS-MED & NowcastHub-MedianEnsemble\\
        \bottomrule
    \end{tabular}
    \caption[Matching the abbreviation to the key in the nowcasting hub.]{Matching the abbreviation to the key in the nowcasting hub.
    Information on the models and references is listed in \citet[][Table 1]{Wolffram2023}.}
    \label{tab:app-covid-models}
\end{table}

\begin{table}
    \centering
    \tiny
    \input{plots/covid_nowcast/10_marginal_analysis}
    \caption[Marginal analysis of the Covid nowcast and true differences for the horizons 1, 7, and 14 days.]{Marginal analysis of the nowcast and true differences for the horizons 1, 7, and 14 days.
    The column (1), $l=l$ shows the number of values greater than zero for horizon $l$, $\sigma_{x^{\Delta, l}}$ the standard deviation, and $q_{0.1} (x^{\Delta, l})$ the 10\% quantile of the differences' absolute values.}
    \label{tab:app-covid-marginals}
\end{table}

\begin{table}
    \centering
    \begin{subtable}[t]{\textwidth}
        \input{plots/covid_nowcast/30_trending_ratios_lag_1}
    \caption{1 day.}
    \end{subtable}
    \begin{subtable}[t]{\textwidth}
        \input{plots/covid_nowcast/30_trending_ratios_lag_14}
        \caption{14 days.}
    \end{subtable}
    \caption[Trending ratios for the models with and without exclusion areas for the horizon 1 and 14 days in Covid nowcasting.]{Trending ratio $\acc$, positive trending ratio $\accp$, and negative trending ratio $\accm$ for the models with and without exclusion areas for the horizon 1 and 14 days. The exclusion areas are rectangles centered on the zero points with a width and height of 10\% of the quantile of the absolute values of nowcast and true values. }
    \label{tab:app-covid-trending-ratios-lag-1-14}
\end{table}

\begin{figure}
\centering
\includegraphics{plots/covid_nowcast/30_4q_plots}
\caption[Four-quadrant plots for the Covid nowcast models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days.]{Four-quadrant plots for the models ILM, RIVM, RKI, and ENS-MEAN and the horizons of one, seven, and 14 days. The spread in both directions increases with the horizon. }
\label{fig:app-covid-4q}
\end{figure}
